---
title: "Data Practical 7"
author: "Stefano Politi"
date: "(`r format(Sys.time(), '%d %B %Y')`)"
output:
  github_document:
    toc: TRUE
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, echo = TRUE)
```

Required packages:

```{r packages}
library(readxl)
library(knitr)
library(dplyr)
library(tidyr)
library(ggplot2)
```

## Importing Data

The following dataset is described [here](https://github.com/StefPo21/IntroDataScience/tree/main/3_data/2_Data.md#Importing-External-Data) (Dataset 2).

```{r GEMMES}
GEMMES <- read_xlsx("Exp2A-GEMMES_Binomial.xlsx", sheet = 1)
```

The dataset contains multiple copies for each participant. For the purpose of this practice, the duplicates will be removed.

```{r GEMMES_clean}
GEMMES_clean <- GEMMES %>%
  select(-VISU, -VISU_Value, -VISU_Binomial)
GEMMES_clean <- unique(GEMMES_clean)
```

The dataset contains many variables. For the purpose of this practice, only a subset will be selected.

```{r GEMMES_selection}
GEMMES2 <- GEMMES_clean %>%
  select(Participant_ID, Sex, Age, Musician_type, Excerpt_id, Excerpt_GEMS, BEST_GEMS, BEST_VA, BEST_VISU)
kable(head(GEMMES2))
```

## Hypothesis

Musical expertise and sex should not influence the relationship between the participant ratings for musical metaphors (GEMMES) and musical emotions (GEMS).

## Filtering Data

There are no participants classified as musician type 1 (no musical experience).
Type 2 corresponds to non musicians, with a minimum of musical experience.
Type 3 corresponds to musicians, with at least 5 years of regular instrument practice.s

```{r musician types}
GEMMES2_types <- GEMMES2 %>%
  count(Musician_type)
kable(GEMMES2_types)
GEMMES2_type2 <- GEMMES2 %>%
  filter(Musician_type == 2)
GEMMES2_table2 <- GEMMES2_type2 %>%
  count(var=GEMMES2_type2$BEST_GEMS, wt_var = GEMMES2_type2$BEST_VISU)
kable(GEMMES2_table2)
GEMMES2_type3 <- GEMMES2 %>%
  filter(Musician_type == 3)
GEMMES2_table3 <- GEMMES2_type3 %>%
  count(var=GEMMES2_type3$BEST_GEMS, wt_var = GEMMES2_type3$BEST_VISU)
kable(GEMMES2_table3)
```

```{r sex}
GEMMES2_sexes <- GEMMES2 %>%
  count(Sex)
kable(GEMMES2_sexes)
GEMMES2_M <- GEMMES2 %>%
  filter(Sex == "M")
GEMMES2_tableM <- GEMMES2_M %>%
  count(var=GEMMES2_M$BEST_GEMS, wt_var = GEMMES2_M$BEST_VISU)
kable(GEMMES2_tableM)
GEMMES2_F <- GEMMES2 %>%
  filter(Sex == "F")
GEMMES2_tableF <- GEMMES2_F %>%
  count(var=GEMMES2_F$BEST_GEMS, wt_var = GEMMES2_F$BEST_VISU)
kable(GEMMES2_tableF)
```

## Plotting Data

The grouped bar-plots for both musician types seem identical.

```{r plot_type}
ggplot(GEMMES2_table2, aes(x=var, y=n, fill = wt_var))+
  geom_bar(position='dodge', stat='identity')+
  theme(axis.text.x = element_text(size = 7))+
  scale_x_discrete(labels = c("Joyful Activation", "Nostalgia", "Peacefulness", "Power", "Sadness", "Tenderness", "Tension"))+
  labs(title = "Frequency of GEMMES ratings for each GEMS ratings (non-musicians)", x="GEMS (ratings)", y="Frequency", fill = "GEMMES (ratings)")+
  scale_fill_discrete(labels = c("Flow", "Force", "Interior", "Movement"))

ggplot(GEMMES2_table3, aes(x=var, y=n, fill = wt_var))+
  geom_bar(position='dodge', stat='identity')+
  theme(axis.text.x = element_text(size = 7))+
  scale_x_discrete(labels = c("Joyful Activation", "Nostalgia", "Peacefulness", "Power", "Sadness", "Tenderness", "Tension"))+
  labs(title = "Frequency of GEMMES ratings for each GEMS ratings (musicians)", x="GEMS (ratings)", y="Frequency", fill = "GEMMES (ratings)")+
  scale_fill_discrete(labels = c("Flow", "Force", "Interior", "Movement"))
```

The grouped bar-plots for both sexes seem identical.

```{r plot_sex}
ggplot(GEMMES2_tableM, aes(x=var, y=n, fill = wt_var))+
  geom_bar(position='dodge', stat='identity')+
  theme(axis.text.x = element_text(size = 7))+
  scale_x_discrete(labels = c("Joyful Activation", "Nostalgia", "Peacefulness", "Power", "Sadness", "Tenderness", "Tension"))+
  labs(title = "Frequency of GEMMES ratings for each GEMS ratings (males)", x="GEMS (ratings)", y="Frequency", fill = "GEMMES (ratings)")+
  scale_fill_discrete(labels = c("Flow", "Force", "Interior", "Movement"))

ggplot(GEMMES2_tableF, aes(x=var, y=n, fill = wt_var))+
  geom_bar(position='dodge', stat='identity')+
  theme(axis.text.x = element_text(size = 7))+
  scale_x_discrete(labels = c("Joyful Activation", "Nostalgia", "Peacefulness", "Power", "Sadness", "Tenderness", "Tension"))+
  labs(title = "Frequency of GEMMES ratings for each GEMS ratings (females)", x="GEMS (ratings)", y="Frequency", fill = "GEMMES (ratings)")+
  scale_fill_discrete(labels = c("Flow", "Force", "Interior", "Movement"))
```

## Analysing Data

Proportion test could be useful to analyse the data, however I did not manage to properly do such tests...

## References

Schaerlaeken, S., Glowinski, D., & Grandjean, D. (2022). Linking musical metaphors and emotions evoked by the sound of classical music. *Psychology of music*, 50(1), 245-264.