---
title: "Data Practical 3"
author: "Stefano Politi"
date: "(`r format(Sys.time(), '%d %B %Y')`)"
output:
  github_document:
    toc: TRUE
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, echo = TRUE)
```

```{r packages}
library(dslabs)
library(readxl)
library(knitr)
library(dplyr)
```
## External Data

### Importing Data

The following dataset is described [here](https://github.com/StefPo21/IntroDataScience/tree/main/3_data/2_Data.md#Importing-External-Data) (Dataset 2).

```{r GEMMES}
GEMMES <- read_xlsx("Exp2A-GEMMES_Binomial.xlsx", sheet = 1)
```

### Filtering Data

The dataset contains multiple copies for each participant. For the purpose of this practice, the duplicates will be removed.

```{r GEMMES_clean}
GEMMES_clean <- GEMMES %>%
  select(-VISU, -VISU_Value, -VISU_Binomial)
GEMMES_clean <- unique(GEMMES_clean)
kable(head(GEMMES_clean, 10))
```

Here the focus will be exclusively on the music excerpts associated to GEMS emotions of "Power" and "Tension" and on their respective highest participant ratings for GEMS, VA and GEMMES.

```{r GEMMES_filter}
GEMMES2 <- GEMMES_clean %>%
  filter(Excerpt_GEMS %in% c("Power", "Tension")) %>%
  select(Participant_ID, Excerpt_GEMS, BEST_GEMS, BEST_VA, BEST_VISU)
kable(head(GEMMES2, 10))
kable(tail(GEMMES2, 10))
```

### Structuring Data

The structure of the filtered dataset shows 1 600 observation of 4 variables.

The variables are listed here with their respective data types:

* BEST_GEMS: character variable (qualitative, categorical) indicating the GEMS emotion with the highest rating for each participant.
* BEST_VA: character variable (qualitative, categorical) indicating highest rating of either valence or arousal for each participant.
* BEST_VISU: character variable (qualitative, categorical) indicating the GEMMES metaphor with the highest rating for each participant.

```{r GEMMES_structure}
str(GEMMES2)
```

When ordering the data by the GEMS emotion to which each music excerpt is associated, there seems to be a strong correlation with the participant's ratings.

```{r GEMMES_ordered}
GEMMES2 <- GEMMES2 %>%
  arrange(Excerpt_GEMS)
kable(head(GEMMES2, 20))
kable(tail(GEMMES2, 20))
```

### Summarizing data

To be able to verify this correlation, the data must be grouped and the occurrences of each rating counted.

The results show indeed a strong correlations:

* "Power" is 100% correlated with ratings of "Power", "Arousal" and "Force".
* "Tension" is 100% correlated with ratings of "Tension", as well as:
  + 50% correlated with ratings of "Arousal" and "Force.
  + 50% correlated with ratings of "Valence" and "Movement".

```{r GEMMES_counts}
GEMMES3 <- GEMMES2 %>%
  group_by(Excerpt_GEMS) %>%
  count(BEST_GEMS, BEST_VA, BEST_VISU)
kable(print(GEMMES3))
```

### Removing NAs

There are no NA values are present in the filtered dataset.

```{r GEMMES_NA}
GEMMES4 <- na.omit(GEMMES2)
str(GEMMES4)
```

## R Pre-loaded Data

### Joining data

The [results_us_election_2016](https://www.rdocumentation.org/packages/dslabs/versions/0.7.4/topics/polls_us_election_2016) dataset, from the dslabs package, describes the poll results of the US 2016 presidential elections.

```{r us_elections}
kable(head(results_us_election_2016))
```

The [murders](https://www.rdocumentation.org/packages/dslabs/versions/0.7.4/topics/murders) dataset, also from the dslabs package, describes gun murder data from FBI reports in the US, organized by state.

```{r murders}
kable(head(murders))
```

Joining the two datasets might be useful to investigate the relationship between the amount of gun murders in a state and the tendency of its citizens to vote for either Trump of Clinton.

```{r joinData}
joinedData <- full_join(results_us_election_2016, murders)
kable(head(joinedData, 10))
```

## References

Schaerlaeken, S., Glowinski, D., & Grandjean, D. (2022). Linking musical metaphors and emotions evoked by the sound of classical music. *Psychology of music*, 50(1), 245-264.