---
title: "Data Practical 5"
author: "Stefano Politi"
date: "(`r format(Sys.time(), '%d %B %Y')`)"
output:
  github_document:
    toc: TRUE
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, echo = TRUE)
```

Required packages:

```{r packages}
library(readxl)
library(knitr)
library(dplyr)
library(tidyr)
library(ggplot2)
```

## Importing Data

The following dataset is described [here](https://github.com/StefPo21/IntroDataScience/tree/main/3_data/2_Data.md#Importing-External-Data) (Dataset 2).

```{r GEMMES}
GEMMES <- read_xlsx("Exp2A-GEMMES_Binomial.xlsx", sheet = 1)
```

The dataset contains multiple copies for each participant. For the purpose of this practice, the duplicates will be removed.

```{r GEMMES_clean}
GEMMES_clean <- GEMMES %>%
  select(-VISU, -VISU_Value, -VISU_Binomial)
GEMMES_clean <- unique(GEMMES_clean)
```

The dataset contains many variables. For the purpose of this practice, only a subset will be selected.

```{r GEMMES_selection}
GEMMES2 <- GEMMES_clean %>%
  select(Sex, Age, Musician_type, Excerpt_id, Excerpt_GEMS, 79:97)
kable(head(GEMMES2, 10))
```

## Transforming Data

Transforming the dataset from wide to long format:

```{r GEMMES_transform}
GEMMES2_long <- gather(GEMMES2, condition, measurement, VISU_Flow:VISU_Wandering)
kable(head(GEMMES2_long, 10))
```

Uniting columns concerning participant information (sex, age and musician type):

```{r GEMMES_unite}
GEMMES2_long <- GEMMES2_long %>%
  unite(Participant_info, Sex, Age, Musician_type)
kable(head(GEMMES2_long, 10))
```

Reverting the previous operation, thus separating columns concerning participant information (sex, age and musician type):

```{r GEMMES_separate}
GEMMES2_long <- GEMMES2_long %>%
    separate(Participant_info, into=c("Sex", "Age", "Musician_type"), sep ="_")
kable(head(GEMMES2_long, 10))
```

## Plotting Data

The aim of this section is to plot participant ratings of musical metaphors (GEMMES) against musical emotions (GEMS) previously attributed to each excerpt.

To do so, the first step is to calculate the average ratings of each metaphor for each emotion.

```{r GEMS_GEMMES_table}
GEMMES2_table <- GEMMES2 %>%
  group_by(Excerpt_GEMS) %>%
  summarize(mean_Flow = mean(VISU_Flow), mean_Force = mean(VISU_Force), mean_Interior = mean(VISU_Interior), mean_Movement = mean(VISU_Movement), mean_Wandering = mean(VISU_Wandering))
GEMMES2_table_long <- gather(GEMMES2_table, condition, measurement, mean_Flow:mean_Wandering)
kable(head(GEMMES2_long, 10))
```

Now, a grouped barplot can be created, visuualizing the average metaphor ratings for each attributed emotion.

```{r GEMS_GEMMES_bar}
ggplot(GEMMES2_table_long, aes(x=Excerpt_GEMS, y=measurement, fill = condition))+
  geom_bar(position='dodge', stat='identity')+
  theme(axis.text.x = element_text(size = 6))+
  scale_x_discrete(labels = c("Joyful Activation", "Nostalgia", "Peacefulness", "Power", "Sadness", "Tenderness", "Tension", "Transcendance", "Wonder"))+
  labs(title = "Average GEMMES ratings for each attributed GEMS emotion", x="GEMS (attributed)", y="Average rating (%)", fill = "GEMMES (ratings)")+
  scale_fill_discrete(labels = c("Flow", "Force", "Interior", "Movement", "Wandering"))
```

## References

Schaerlaeken, S., Glowinski, D., & Grandjean, D. (2022). Linking musical metaphors and emotions evoked by the sound of classical music. *Psychology of music*, 50(1), 245-264.