---
title: "Data Practical 6"
author: "Stefano Politi"
date: "(`r format(Sys.time(), '%d %B %Y')`)"
output:
  github_document:
    toc: TRUE
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, echo = TRUE)
```

Required packages:

```{r packages}
library(readxl)
library(knitr)
library(dplyr)
library(tidyr)
library(ggplot2)
```

## Importing Data

The following dataset is described [here](https://github.com/StefPo21/IntroDataScience/tree/main/3_data/2_Data.md#Importing-External-Data) (Dataset 2).

```{r GEMMES}
GEMMES <- read_xlsx("Exp2A-GEMMES_Binomial.xlsx", sheet = 1)
```

The dataset contains multiple copies for each participant. For the purpose of this practice, the duplicates will be removed.

```{r GEMMES_clean}
GEMMES_clean <- GEMMES %>%
  select(-VISU, -VISU_Value, -VISU_Binomial)
GEMMES_clean <- unique(GEMMES_clean)
```

The dataset contains many variables. For the purpose of this practice, only a subset will be selected.

```{r GEMMES_selection}
GEMMES2 <- GEMMES_clean %>%
  select(Sex, Age, Musician_type, Excerpt_id, Excerpt_GEMS, BEST_GEMS, BEST_VA, BEST_VISU)
kable(head(GEMMES2, 10))
```

## Research questions

Did the musical emotions (GEMS) which were pre-attributed to each musical excerpt correspond to the participant's ratings of the musical emotions (GEMS)?

Which musical metaphors (GEMMES) ratings are best correlated with either the attributed or the rated musical emotions (GEMMES)?

## Plotting Data

The plot of musical emotions (GEMS) ratings for each attributed musical emotion (GEMS) shows the following correlations:

* Joyful Activation, Nostalgia, Power, Tension and Wonder ratings show a frequency of 100% for their respective prior attributions.
* Peacefulness, Sadness, Tenderness and Transcendence show a frequency of 50% for their respective prior attributions.
  + Peacefulness shows a frequency of 50% also for Tenderness ratings.
  + Sadness, Tenderness and Transcendence show a frequency of 50% also for Nostalgia ratings.
  
Note: The ratings used for the following visualizations are only the highest ratings of each participant for each music excerpt. They do not take into consideration all other participant's ratings (which might be very close to the best). For example an excerpt might be rated, on a scale from 0 to 100, as 65 Sadness and 64 Nostalgia, but only Sadness will be counted as most representative of the excerpt for this participant.
  
```{r GEMS}
GEMMES2_table <- GEMMES2 %>%
  count(var=GEMMES2$Excerpt_GEMS, wt_var = GEMMES2$BEST_GEMS)
ggplot(GEMMES2_table, aes(x=var, y=n, fill = wt_var))+
  geom_bar(position='dodge', stat='identity')+
  theme(axis.text.x = element_text(size = 6))+
  scale_x_discrete(labels = c("Joyful Activation", "Nostalgia", "Peacefulness", "Power", "Sadness", "Tenderness", "Tension", "Transcendance", "Wonder"))+
  labs(title = "Frequency of GEMS ratings for each attributed GEMS musical emotion", x="GEMS (attributed)", y="Frequency", fill = "GEMS (ratings)")+
  scale_fill_discrete(labels = c("Joyful Activation", "Nostalgia", "Peacefulness", "Power", "Sadness", "Tenderness", "Tension", "Transcendance", "Wonder"))
```
The plot of musical metaphors (GEMMES) ratings for each attributed musical emotion (GEMS) shows the following correlations:

* Joyful Activation and Wonder show a frequency of 100% for Movement.
* Peacefulness and Tenderness show a frequency of 100% for Flow.
* Power shows a frequency of 100% for Force.
* Nostalgia and Sadness show a 50% frequency for Flow and 50% for Interior.
* Tension shows a 50% frequency for Force and 50% for Movement.
* Transcendence shows a 50% frequency for Flow and 50% for Force.

Note: Wandering is not represented as best GEMMES rating for any excerpt.

```{r GEMSattr_GEMMES}
GEMMES2_table <- GEMMES2 %>%
  count(var=GEMMES2$Excerpt_GEMS, wt_var = GEMMES2$BEST_VISU)
ggplot(GEMMES2_table, aes(x=var, y=n, fill = wt_var))+
  geom_bar(position='dodge', stat='identity')+
  theme(axis.text.x = element_text(size = 6))+
  scale_x_discrete(labels = c("Joyful Activation", "Nostalgia", "Peacefulness", "Power", "Sadness", "Tenderness", "Tension", "Transcendance", "Wonder"))+
  labs(title = "Frequency of GEMMES ratings for each attributed GEMS musical emotion", x="GEMS (attributed)", y="Frequency", fill = "GEMMES (ratings)")+
  scale_fill_discrete(labels = c("Flow", "Force", "Interior", "Movement"))
```
The plot of musical metaphors (GEMMES) ratings for musical emotion (GEMS) ratings shows the following correlations:

* Joyful Activation shows a frequency of 100% for Movement (idem as previously, for attributed GEMS).
* Peacefulness and Tenderness shows a frequency of 100% for Flow (idem as previously, for attributed GEMS).
* Tension shows a 50% frequency for Force and 50% for Movement (idem as previously, for attributed GEMS).
* Sadness shows a 100% frequency for Interior (unlike previously, which showed 50% Interior and 50% Flow).
* Power shows a slightly higher frequency for Force than for Movement (unlike previously, which showed 100% Force).
* Nostalgia shows a higher frequency for Flow than for Interior (unlike previously, which showed 50% for each).

Note: Transcendence and Wonder are not represented as best GEMS ratings for any excerpt.

```{r GEMSratings_GEMMES}
GEMMES2_table <- GEMMES2 %>%
  count(var=GEMMES2$BEST_GEMS, wt_var = GEMMES2$BEST_VISU)
ggplot(GEMMES2_table, aes(x=var, y=n, fill = wt_var))+
  geom_bar(position='dodge', stat='identity')+
  theme(axis.text.x = element_text(size = 7))+
  scale_x_discrete(labels = c("Joyful Activation", "Nostalgia", "Peacefulness", "Power", "Sadness", "Tenderness", "Tension"))+
  labs(title = "Frequency of GEMMES ratings for each GEMS ratings", x="GEMS (ratings)", y="Frequency", fill = "GEMMES (ratings)")+
  scale_fill_discrete(labels = c("Flow", "Force", "Interior", "Movement"))
```

## References

Schaerlaeken, S., Glowinski, D., & Grandjean, D. (2022). Linking musical metaphors and emotions evoked by the sound of classical music. *Psychology of music*, 50(1), 245-264.