---
title: "Data Practical 8"
author: "Stefano Politi"
date: "(`r format(Sys.time(), '%d %B %Y')`)"
output:
  github_document:
    toc: TRUE
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, echo = TRUE)
```

Required packages:

```{r packages}
library(tidyr)
library(knitr)
library(dslabs)
library(ggplot2)
```

## Importing Data

The [cars](https://www.rdocumentation.org/packages/datasets/versions/3.6.2/topics/cars) dataset shows the speed of the cars and distances taken to stop (note that the data were recorded in the 1920).

```{r cars}
data("cars")
kable(head(cars))
```

The dataset is composed of 50 observations for 2 variables:
* Explanatory variable: speed (numeric: quantitative, continuous).
* Response variable: dist (numeric: quantitative, continuous).

```{r cars_structure}
str(cars)
```
## Linear model

There is a high correlation between the 2 variables (r = 0.8) and the p-value is significant (p = 0.01).

```{r model}
cor.test(cars$speed, cars$dist)
model<-lm( dist ~speed, data= cars)
summary(model)
```

## Plot

The plot clearly shows an increase in the distance to stop as the speed increases.

```{r plot}
ggplot(cars, aes(speed, dist)) +
  geom_point() +
  geom_smooth(method = "lm")
```